---
title: "Class 09"
format: pdf
editor: visual
---

## PDB Exploration

We will start by load a csv file from PDB

To read this file we are going to use read.csv

```{r}
library(readr)
pdb_stats <- read_csv("Data Export Summary.csv")
pdb_stats
```

We are going to explore the data

```{r}
pdb_stats$X.ray
```

We are going to use gsub to remove commas

```{r}
as.numeric(gsub(",", "", pdb_stats$X.ray))
```

```{r}
as.numeric(gsub(",", "", pdb_stats$EM))
```

I use sum command to get sum

```{r}
n_xray = sum(as.numeric(gsub(",", "", pdb_stats$X.ray)))
```

```{r}
n_em = sum(as.numeric(gsub(",", "", pdb_stats$EM)))
```

```{r}
n_total = sum(as.numeric(gsub(",", "", pdb_stats$Total)))
```

Q1. What percentage of structures in the PDB are solved by X-Ray and Electron Microscopy.

```{r}
((n_xray + n_em)/n_total) * 100
```

The percentage of structures solved by X-Ray and EM is 92.99%

**Q2:** What proportion of structures in the PDB are protein?

```{r}
n_proteino = as.numeric(gsub(",", "", pdb_stats["Protein (only)","Total"]))
n_proteino/n_total
```

0.86 proportion of the structures are proteins

```{r}
sum(as.numeric(gsub(",", "", pdb_stats$Total)))
```

Q3. Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

Too difficult

# 2. Visualizing the HIV-1 protease structure

![](1HSG%20(1).png)

There is a critical "conserved" water molecule in the binding site. Can you identify this water molecule? What residue number does this water molecule have

Residue 308

![](1HSG%20(2).png)

# Intro to BIO3d in R

```{r}
library(bio3d)
pdb= read.pdb("1hsg")
```

```{r}
pdb
```

```{r}
attributes(pdb)
```

```{r}
head(pdb$atom)
```

# Predicting functional motions of a single structure by NMA

```{r}
adk = read.pdb("6s36")
```

```{r}
adk
```

```{r}
m = nma(adk)
```

```{r}
plot(m)
```

```{r}
mktrj(m, file = "adk_m7 .pdb")
```
