---
title: "HW 06"
author: "Arturo Agramont"
date: "2023-04-25"
output:
  html_document: default
  pdf: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Protein Code

First install and load the bio3d package

```{r}
#install.packages("bio3d")
library(bio3d)
```

I will begin by running the code provided

```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

I will start by trying to apply all portions to "4AKE" or "s1"

Firstly starting with trim in the function

```{r}
trimpro = function(x){ #naming the function for trimming protein
  trim.pdb(x, chain="A", elety="CA") #using trim function on pdb input
}
trimpro(s1)
```

The function is providing a trim of each pdb file.

```{r}
trimpro(s2)
```

Now we will try to implement the "\$atom\$b" into the function

```{r}
atom = function(x){ #naming the function protein
  trimpro(x)$atom$b #adding the atom portion
}
atom(s1)
```

The function provides the atom numbers.

Now we will try to add the plot portion to the function

```{r}
#' Protein function to provide a plot
#'
#' @param x, or input is a loaded pdb file. For Example: s1 is a loaded file from bio3d using function read.pdb()
#'
#' @return returns a plot for specified protein
#'
#'
#' @examples
protein = function(x){ #naming the function protein
  plotb3(atom(x), sse = trimpro(x), typ="l", ylab="Bfactor") #implementing both functions I had previously written in order to make a plot 
}
protein(s1)
```

Now it will test on the rest of the proteins

```{r}
protein(s2)

```

```{r}
protein(s3)
```
