---
title: "Class 10"
author: Arturo Agramont
format: pdf
editor: visual
---

# Importing candy data

```{r}
candy = read.csv("candy-data.csv", row.names = 1)
head(candy)
```

-   **Q1**. How many different candy types are in this dataset?

```{r}
nrow(candy)
```

There are 85 kinds of candy in this dataset.

-   **Q2**. How many fruity candy types are in the dataset?

```{r}
table(candy$fruity)
```

There are 38 fruity candy types in the data set.

# Viewing win percent

```{r}
candy["Twix", ]$winpercent
```

Twix in this case has an 81.64% win rate in match-ups against other candies.

-   **Q3**. What is your favorite candy in the dataset and what is it\'s `winpercent` value?

```{r}
candy["Sour Patch Kids",]$winpercent
```

The win percent value for my favorite candy, Sour Patch Kids, is 59.86%

-   **Q4**. What is the `winpercent` value for \"Kit Kat\"?

```{r}
candy["Kit Kat",]$winpercent
```

Kit Kat has a 76.76% win rate

-   **Q5**. What is the `winpercent` value for \"Tootsie Roll Snack Bars\"?

```{r}
candy["Tootsie Roll Snack Bars",]$winpercent
```

Tootsie rolls have a 49.65% win percent value.

We are going to use skim to look at the data

```{r}
library("skimr")
skim(candy)
```

-   **Q6**. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

It appears that win percent is on a different scale than the rest of the columns.

-   **Q7**. What do you think a zero and one represent for the `candy$chocolate` column?

A zero represents that the candy has no chocolate and a 1 is that the candy contains chocolate.

We will now start making plots

-   **Q8**. Plot a histogram of `winpercent` values

```{r}
hist(candy$winpercent)
```

-   **Q9**. Is the distribution of `winpercent` values symmetrical?

The distribution is not symmetrical but is shifted to the left.

-   **Q10**. Is the center of the distribution above or below 50%?

Below 50%

-   **Q11**. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
mean(candy$winpercent[as.logical(candy$fruity)])

```

```{r}
mean(candy$winpercent[as.logical(candy$chocolate)])
```

On average chocolate is chosen more and higher ranked.

-   **Q12**. Is this difference statistically significant?

```{r}
t.test(candy$winpercent[as.logical(candy$chocolate)],
       candy$winpercent[as.logical(candy$fruity)])
```

This is a significant difference.

# Overall candy rankings

-   **Q13**. What are the five least liked candy types in this set?

    We will use order() to arrange the data set by win percent

```{r}
head(candy[order(candy$winpercent),])


```

The 5 least likes are Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, Jawbusters

-   **Q14**. What are the top 5 all time favorite candy types out of this set?

```{r}
tail(candy[order(candy$winpercent),])
```

Reese's Peanut Butter Cup, Reese's Miniatures, Twix, Kit Kat, and Snickers have the highest win percent.

-   **Q15**. Make a first barplot of candy ranking based on `winpercent` values.

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```

-   **Q16**. This is quite ugly, use the `reorder()` function to get the bars sorted by `winpercent`?

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col()
```

Now we are going to add color to the graph

We will assign colors to a variable to apply to the bar graph

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols)
```

\- **Q17**. What is the worst ranked chocolate candy?

Sixlets are the worst ranked chocolate candy.

\
- **Q18**. What is the best ranked fruity candy?

Starburst is the best ranked fruity candy

# Price percent

We are going to make a plot of win percent vs price percent, we will use ggplot and ggrepel

```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```

-   **Q19**. Which candy type is the highest ranked in terms of `winpercent` for the least money - i.e. offers the most bang for your buck?

Reese's miniatures, starburst, kit kat are the highest ranked in win percent for the least amount of money

-   **Q20**. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

We are going to change the order of the data in order to see the most expensive candy and the least popular of the 5 most expensive.

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )

```

The least popular expensive candy is Nik L Nip.

# Correlation Structure

We are going to use corrplot to explore correlation

```{r}
library(corrplot)
```

```{r}
cij <- cor(candy)
corrplot(cij)
```

-   **Q22**. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

Win percent and price percent are the most anti-correlated

-   **Q23**. Similarly, what two variables are most positively correlated?

Chocolate and Fruity are the most positive correlated

# Using PCA 

We will use PCA analysis on the data

```{r}
pca = prcomp(candy, scale = T)
summary(pca)
```

Now we will plot PC1 vs PC2

```{r}
plot(pca$x[,1:2])
```

Now adding color

```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```
