---
title: "Class 17"
author: Arturo Agramont
format: pdf
editor: visual
---

# Covid-19 Vaccine data

We will start by loading the data

```{r}
vax = read.csv("covid19.csv")
head(vax)
```

-   **Q1.** What column details the total number of people fully vaccinated?

The column that details the number of people fully vaccinated is: persons_fully_vaccinated

-   **Q2.** What column details the Zip code tabulation area?

the column that details this information is zip_code_tabulation_area

-   **Q3.** What is the earliest date in this dataset?

The earliest date in this data set is January 5th, 2021.

-   **Q4.** What is the latest date in this dataset?

```{r}
tail(vax)
```

The latest data in this data set is June 6th, 2023

We will now call skimr for an overview of the data

```{r}
skimr::skim_without_charts(vax)
```

**Q5.** How many numeric columns are in this dataset?

There are 14 numeric columns in this data set.

**Q6.** Note that there are \"missing values\" in the dataset. How many `NA` values there in the `persons_fully_vaccinated` column?

```{r}
sum(is.na(vax$persons_fully_vaccinated))
```

There are 17966 NA values

**Q7.** What percent of `persons_fully_vaccinated` values are missing (to 2 significant figures)?

```{r}
((sum(is.na(vax$persons_fully_vaccinated)))/nrow(vax)) * 100


```

8.0% of persons_fully_vaccinated values are missing

-   **Q8.** \[Optional\]: Why might this data be missing?

This could be because not all of the patients or people in the study are fully vaccinated.

# Working with dates

Using lubridate

```{r}
library(lubridate)
```

```{r}
today()
```

Converting vax column of dates into year month day format in order to plot

```{r}
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
today() - vax$as_of_date[1]
```

Then we can determine the days between the data set

```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

**Q9.** How many days have passed since the last update of the dataset?

```{r}
table(vax$as_of_date)
```

7 days have passed between updates.\
**Q10.** How many unique dates are in the dataset (i.e. how many different dates are detailed)?

```{r}
library(dplyr)
n_distinct(vax$as_of_date)

```

There are 127 unique dates in this data set

# Working with zip codes

We are going to use zipcodeR package

```{r}
library(zipcodeR)

```

```{r}
geocode_zip('92037')

```

calculating distance between zip codes

```{r}
zip_distance('92037','92109')

```

Collecting census data of the zip codes:

```{r}
reverse_zipcode(c('92037', "92109") )

```

Focusing on San Diego

```{r}
sd = vax[vax$county == "San Diego",]
head(sd)
```

**Q11.** How many distinct zip codes are listed for San Diego County?

```{r}
n_distinct(sd$zip_code_tabulation_area)
```

There are 107 distinct zip code areas

**Q12.** What San Diego County Zip code area has the largest population in this dataset?

```{r}
sd.80 <- filter(vax, county == "San Diego" &
                tot_population > 80000)
head(sd.80)
```

Zip code 92154 has the highest total population

**Q13.** What is the overall average (with 2 decimal numbers) \"Percent of Population Fully Vaccinated\" value for all San Diego \"County\" as of \"2023-05-23\"?

```{r}
sd2 <- filter(vax, county == "San Diego" & as_of_date == "2023-05-23")
mean(na.omit(sd2$percent_of_population_fully_vaccinated))*100
```

74.20% is the average of percent of population fully vaccinated

-   **Q14.** Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of \"2023-05-23\"?

```{r}
library(ggplot2)
ggplot(sd2)  + geom_histogram(aes(percent_of_population_fully_vaccinated), bins = 9)
```

# UCSD/ La Jolla

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```

-   **Q15**. Using **ggplot** make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
ggplot(ucsd) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")

```

Looking at data of larger sized populations

```{r}
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2023-05-23")


```

-   **Q16**. Calculate the mean *\"Percent of Population Fully Vaccinated\"* for ZIP code areas with a population as large as 92037 (La Jolla) *as_of_date* \"2023-05-23\". Add this as a straight horizontal line to your plot from above with the `geom_hline()` function?

```{r}
mean(vax.36$percent_of_population_fully_vaccinated)
```

```{r}
ggplot(ucsd) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated") +
  geom_hline(yintercept = mean(vax.36$percent_of_population_fully_vaccinated))
```

-   **Q17.** What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the *\"Percent of Population Fully Vaccinated\"* values for ZIP code areas with a population as large as 92037 (La Jolla) *as_of_date* \"2023-05-23\"?

```{r}
summary(vax.36)
```

-   **Q18.** Using ggplot generate a histogram of this data.

```{r}
ggplot(vax.36)+ geom_histogram(aes(percent_of_population_fully_vaccinated), bins = 17)
```

-   **Q19**. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
vax %>% filter(as_of_date == "2023-05-23") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```

-   **Q20.** Finally make a time course plot of vaccination progress for all areas in the full dataset with a `age5_plus_population > 36144`.

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(0,1) +
  labs(x="Date", y="Percent Vaccinated",
       title="Vaccination rate across California",
       subtitle="only areas above 36k population") +
  geom_hline(yintercept = mean(vax.36.all$percent_of_population_fully_vaccinated), linetype= "dash")
```
